---
title: "ESM 206 Lab 10"
author: "Sam Csik"
date: "December 3, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lab 10 Objectives:

- Correlation matrices
- Binary logistic regression
- Dates/times with lubridate
- Heatmap of solar irradiation in Santa Barbara, Hilo (HI), and Fairbanks (AK) in 2010
- Aquaculture with gganimate

# PART 0. Load libraries
```{r packages}

library(tidyverse) # The tidyverse!
library(corrplot) # For correlation matrices
library(janitor) # For cleaning up column names
library(lubridate) # For dealing with dates & times
library(gganimate) # For adding animation to ggplot graphics
library(ggimage) # For updating graph components with images
```

# PART 1. Correlation matrix (World Bank world environmental factors data)

Compiled World Bank data, accessed from: https://www.kaggle.com/zanderventer/environmental-variables-for-world-countries#World_countries_env_vars.csv

- Read in the "world_env_vars.csv file" as env_var
- Find correlations between all variables, columns 2:28
- Plot the correlation matrix (corrplot)

```{r env_corr}
# read in the data
env_var <- read_csv("world_env_vars.csv") %>% 
  na.omit

# calculate Pearson's r for correlations
cor_df <- cor(env_var[2:28])

# make a correlation matrix
corrplot(cor_df,
         type = "upper",
         method = "ellipse", 
         tl.col = "black", 
         tl.cex = 0.5)
# size indicates strength of corrlation
# directionaliity (+, -) indicated by color
  # large dark blue = +
  # small light red = small neg correaltion

# might be concernted about rain_wetestMonth x rain+wetestQuarter (duplicate info)
```

# PART 2. Binary Logistic Regression (Donner Party Data)

Use the 'glm' function for fitting *generalized linear models* (the *logit* - log odds of survival, in our case, will be linearly related to Sex and Age. So we expect the final model to look something like this: 

$$Log Odds (Survival) = \beta_0 + \beta_1(Age) + \beta_2(Sex)$$
~Interpredting coefficients can be a complicated in terms of log-odds~ will see later on how to present these in a way that's easier to understand

We'll use 'family = binomial' to run binomial logistic regression...otherwise, this looks very similar to other types of regression we've already done. 

a. Read in the DonnerTable.csv file as DonnerTable

```{r donner_data}

DonnerTable <- read_csv("DonnerTable.csv")

# outcome of 0 is not-survival; 1 is survival
# Hint: for binary dependent var (T/F, Y/N), specify (using case_when) Y = 0, N = 1 to avoid confusion so that you know both you and R are specifiying 0 as 0 and 1 as 1

```

b. Binomial logistic regression

```{r blr}

donner_blr <- glm(Outcome ~ Sex + Age, family = "binomial", data = DonnerTable)
summary(donner_blr)

# outcome = dependent variable (0 = dead, 1 = survive because Alison told us...but make sure you know which is which when using other data)
# need to specify 

# look at summary (still in log odds)
  # both coeffs for sexMale and Age are negative (even though those are log odds, male    survival was lower, and an increase of Age = decrease in likelihood of survival)

```

c. Use the model to answer: What are the log odds of survival for a 5 year-old female? The probability of survival?

```{r predict_f_5}

# 1. Create a data frame with variables Sex and Age, containing data "Female" and 5, respectively: 

# need new df to make predictions with exact same vars that the model was created with
f_5 <- data.frame(Sex = "Female",
                  Age = 5)

# 2. Find the log odds of survival for the new data (5 year old female) using predict() function with type = "link":

f_5_logodds <- predict(donner_blr, newdata = f_5, type = "link") # using exisiting log-linear equation to make prediction
f_5_logodds # log odds of survival (still challenging to interpret...so convert to odds)

# 3. Exponentiate the log odds to find ODDS of survival for a 5 year old female:

f_5_odds <- exp(f_5_logodds) # 4.23666
# ~4:1 odds of survival; probabilty of survival = 80% (prob happening/prob not happening)

# Ask: Could we manually find the probability of survival for a 5-year old female? recall: p/(1-p) = ODDS

# 4. Actually, let's just use type = "response" in the predict function, which converts to a probability for us:

f_5_prob <- predict(donner_blr, newdata = f_5, type = "response") 
f_5_prob #80.9% probability of survival for age 5 female
```

d. What is the probability of survival for a 25 year-old male? 

```{r predict_m_25}

# Similarly:

m_25 <- data.frame(Sex = "Male", Age = 25) # Make a new data frame

m_25_prob <- predict(donner_blr, newdata = m_25, type = "response") # Find probability of survival
m_25_prob #0.4167068
```

e. Create new sequences of data so that we can graph probabilities for the entire spectrum of ages, designated by sex.

```{r new_data}

seq_age <- rep(seq(from = 0, to = 100), 2) # Create a sequence from 0 to 100, twice (one will be "Male" and one will be "Female")

f_101 <- rep("Female", 101) # Repeat 'Female' 101 times (to match years data)
m_101 <- rep("Male", 101) # Repeat 'Male' 101 times
mf_101 <- c(f_101, m_101) # Combine them into a single vector

# Combine the age and sex sequences into a single data frame - that will be the new data that we have our model make predictions for

donner_newdata <- data.frame(seq_age, mf_101) # MUST make column names match variables in the model!
colnames(donner_newdata) <- c("Age","Sex")

```

f. Now that we have new data to put into our model to have it make predictions, let's go ahead and actually find the predicted probabilities for each Age/Sex combination.

```{r donner_predict}

# Find probabilities using predict (with type = "response"). Include SE.

predicted_probs <- predict(donner_blr, newdata = donner_newdata, type = "response", se.fit = TRUE)

# Coerce outcome into data frame. 

graph_data <- data.frame(donner_newdata, predicted_probs$fit, predicted_probs$se.fit)
colnames(graph_data) <- c("Age", "Sex", "Probability", "SE")
```

g. Graph results.

```{r donner_graph}

ggplot(graph_data, aes(x = Age, y = Probability)) +
  geom_line(aes(color = Sex)) +
  geom_ribbon(aes(ymin = Probability - SE, ymax = Probability + SE, fill = Sex), alpha = 0.4) # lines show predicted prob of survival and ribbons show the SE (+/- 1 SE and each x value)
```


# PART 3. Solar irradiation at the 3 locations in 2010 (pull in all together, do some lubridate stuff, etc.)

a. Read in multiple solar irradiation files (for SB (CA), Hilo (HI), and Fairbanks (AK)):
```{r solar_data}


si_full <- list.files(pattern = "solar_irradiation_*") %>% # * anything after can differ but must have first component before
  map_df(~read_csv(.)) %>% # put together in single df; . tells pipe to use data from immediately before
# site column acts as identifyer
  clean_names()
```

b. Wrangle the data
```{r wrangle_solar}

solar_tidy <- si_full %>% 
  rename(sol_rad = etr_wh_m_2,
         date = yyyy_mm_dd,
         time = hh_mm_lst) %>% 
  filter(time != "NA") %>% 
  mutate(site = fct_relevel(site, "Hawaii", "Santa Barbara", "Alaska")) # can reorder factor levels

```

c. Use lubridate() functions to convert to times/dates
```{r date_time}

solar_tidy$date <- mdy(solar_tidy$date)
solar_tidy$time <- hms(solar_tidy$time)

```

d. Make an awesome figure of solar irradiation (heat/tile map)
```{r solar_graph}

solar_gg <- ggplot(solar_tidy, aes(x = date, y = time)) +
  geom_tile(aes(fill = sol_rad)) + 
  scale_fill_gradientn(colors = c("royalblue2", "mediumorchid1", "orange", "yellow")) + # lets you set as many colors as you want in gradient scheme
  scale_y_time() + 
  facet_wrap(~site, ncol = 1)

solar_gg
```


#PART 4. gganimate example: total aquaculture production (metric tons) for United States, Brazil, Chile, and Ecuador

a. Get the data, and do some wrangling:
```{r aqua_data}

aq_df <- read_csv("aq_wb.csv")

aq_tidy <- aq_df %>% 
  filter(country == "Brazil" |
           country == "Chile" |
           country == "Ecuador" |
           country == "United States") %>% 
  gather(year, aq_prod, `1960`:`2016`) %>% # create new columns year, aq_prod; will store info from original df
  filter(year >= 1990) %>% 
  mutate(aq_mil = aq_prod/1e6) %>% 
  select(country, year, aq_mil)
```


b. Read in the fish.png as 'fish'

```{r fish_image}

fish <- "fish.png"
```

c. Make a graph...with gganimate!
```{r gganimate_fish}

aq_plot <- ggplot(aq_tidy, aes(x = as.numeric(year), y = aq_mil, group = country)) +
  geom_line(aes(color = country)) +
  geom_image(aes(image = fish)) +
  geom_text(aes(label = country, color = country), position = position_nudge(y = 0.4, x = 1), size = 5) +
  transition_reveal(country, as.numeric(year))

aq_plot

# export at a gif
animate(aq_plot, nframe = 24, renderer = gifski_renderer("aq_animate.gif"))

# animation maintained in html
```

##END LAB